Drought vulnerability of tree growth

### 2024-07-07 scripts/explore_analysis.Rmd
visualising top 4 species after workshop meeting

#reading the GitHub PAT (stored privately and locally)
pat<-as.character(read.table("../data/HKK-dendro/git_pat.txt")[1,1])

@@

load("../data/HKK-dendro/DendroByBand.RData")

@@

#this can be refined if needed
#wet_census<-c(1, 3, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28)

@@

chirps_hkk<-read.csv("../data/climate/CHIRPS_from_GEE.csv")
colnames(chirps_hkk)[1]<-"time"

@@

spei_hkk<-read.csv("../data/climate/SPEI_HKK_from_GEE.csv")
colnames(spei_hkk)[1]<-"time"

@@

pre_yr_census<-ggplot()+
  #geom_area(data=chirps_hkk, aes(x=time, y=precipitation), fill="blue", alpha=0.5)+
  geom_line(data=chirps_hkk, aes(x=time, y=precipitation), col="blue", alpha=0.5)+
  geom_line(data=spei_hkk, aes(x=time, y=(100*SPEI_01_month-250)), col="red")+
  geom_rect(data=census_int, aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf), 
              fill = "grey10", alpha = 0.5) +
  geom_text(data=census_int, aes(x=start, y=500, 
                                 label= paste0("C", 1:29), angle=90), size=2)+
   scale_y_continuous(    # Features of the first axis
     name = "monthly rainfall (mm)",

@@

png("../doc/display/census_precip.png", width=6, height=3, res=300, units="in") 
pre_yr_census
dev.off()

@@

dendro.data<-dendrobyband %>%
  dplyr::rename(sp=SPCODE.UPDATE)%>%
  mutate(year=year(Date2)) %>%
  #filter(year!=2209)%>% #quick fix to remove bad dates - there are 30 entries
  #filter(start_date <= year & year <= end_date) %>%
  #filter(Cno %in% wet_census)%>%
  filter(is.na(DBH1_flag), DBH1_error=="none",
         win.decimal.error=="none", is.na(win.flag))%>%
  group_by(Cno, sp, Tag) %>%
  dplyr::summarise(calcDBH=calcDBH[which(n.dendro==min(n.dendro))],
                   Date2=Date2[which(n.dendro==min(n.dendro))],
                   Cii=max(Cii, na.rm=T))

@@

```
```{r calculating increments for specific years}

dendro_2010<- dendro.data.noouts %>%
  filter(Cno %in% c(1, 3, 5))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         #inc=ifelse(inc.time==1, inc, NA),
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

@@

library(gridExtra)
png("../doc/display/increments_raw.png", width=8, height=4, units="in", res=300)
grid.arrange(dendro_2010_plot, dendro_2015_plot, dendro_2019_plot, nrow=1)
dev.off()

@@

```{r long-term growth rates}

library(ggplot2)

pat<-as.character(read.table("../data/HKK-dendro/git_pat.txt")[1,1])

remotes:::download("../data/HKK-dendro/stem1.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem1.rdata",
                   auth_token = pat)
remotes:::download("../data/HKK-dendro/stem2.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem2.rdata",
                   auth_token = pat)
remotes:::download("../data/HKK-dendro/stem3.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem3.rdata",
                   auth_token = pat)
remotes:::download("../data/HKK-dendro/stem4.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem4.rdata",
                   auth_token = pat)

load("../data/HKK-dendro/stem1.RData")
load("../data/HKK-dendro/stem2.RData")
load("../data/HKK-dendro/stem3.RData")
load("../data/HKK-dendro/stem4.RData")

hkk_census <- hkk_census %>%
  filter(status=="A")%>%
  filter(tag %in% dendro.data.noouts$Tag)%>%
  group_by(StemTag)%>%
  dplyr::mutate(dbhmin1 = dplyr::lag(dbh, n=1, default=NA, order_by=StemTag),
                inc_annual = (dbh-dbhmin1)/5)%>%
  filter(!is.na(dbhmin1))%>%
  dplyr::summarise(avg_inc = mean(inc_annual, na.rm=T),
                   sd_inc = sd(inc_annual, na.rm=T))

png("../doc/display/census_increment.png", width=4, height=4, units="in", res=300)
inc_census
dev.off()

@@

@@

resistance_2010 <- dendro_2010 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

@@

resistance_2010_div<-ggplot(data = resistance_2010, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  theme_bw()

@@

resistance_2015 <- dendro_2015 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

@@

resistance_2015_div<-ggplot(data = resistance_2015, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  theme_bw()

@@

@@

resistance_all<-rbind(resistance_2010, resistance_2015, resistance_2019)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010)), rep(2015, nrow(resistance_2015)), rep(2019, nrow(resistance_2019)))

@@

resistance_map<-ggplot(resistance_all, aes(x=X, y=Y, size=as.numeric(DBH.04), col=quantile))+
  geom_point(alpha=0.5)+
   scale_size(range = c(.1, 10), name="DBH (cm)")+
  guides(color=guide_legend(title="resistance"))+
  scale_colour_gradient(low="red", high="blue") +
  facet_grid(yr~.)+
  theme_minimal()+
  coord_fixed()

png("../doc/display/resistance_map.png", width=8, height=10, units="in", res=300)
resistance_map
dev.off()
```
```{r basic model}

model_2010<-glm(resistance.dif ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                data=resistance_2010)

@@

model_2015<-glm(resistance.dif ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                data=resistance_2015)

@@

model_2019<-glm(resistance.dif ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                data=resistance_2019)

@@

resistance_all<-rbind(resistance_2010, resistance_2015, resistance_2019)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010)), rep(2015, nrow(resistance_2015)), rep(2019, nrow(resistance_2019)))

resistance_model<-resistance_all %>%
  select(resistance.dif, calcDBH_min1, cii_min1, williams_dec, yr, quantile)

@@

library(lme4)
library(nlme)
model_all<-lme(resistance.dif ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                random=~1|yr,
                data=resistance_model)

@@

plot_model(model_all, type="pred", terms=c("calcDBH_min1","williams_dec"), show.data = T)
plot_model(model_all, type="pred", terms=c("cii_min1"), show.data = T)
plot_model(model_all, type="pred", terms=c("williams_dec"), show.data = T)
plot_model(model_all, type="pred", terms=c("calcDBH_min1"), show.data = T)

@@

```{r echo=F}

>>>>

#reading the GitHub PAT (stored privately and locally)
#pat<-as.character(read.table("../data/HKK-dendro/git_pat.txt")[1,1])

@@

load("growth-precip-thailand/data/HKK-dendro/DendroByBand.RData")

@@

@@

#chirps_hkk<-read.csv("../data/climate/CHIRPS_from_GEE.csv")
chirps_hkk<-read.csv("growth-precip-thailand/data/climate/CHIRPS_from_GEE.csv")
colnames(chirps_hkk)[1]<-"time"

@@

spei_hkk<-read.csv("growth-precip-thailand/data/climate/SPEI_HKK_from_GEE.csv")
colnames(spei_hkk)[1]<-"time"

@@

pre_yr_census<-ggplot()+
  geom_line(data=chirps_hkk, aes(x=time, y=precipitation), col="blue", alpha=0.5)+
  geom_line(data=spei_hkk, aes(x=time, y=(100*SPEI_01_month-250)), col="red")+
  geom_rect(data=census_int, aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf), 
              fill = "grey10", alpha = 0.5) +
  geom_text(data=census_int, aes(x=start, y=500, 
                                 label= paste0("C", 1:29), angle=90), size=2)+
   scale_y_continuous(    # Features of the first axis
     name = "monthly rainfall (mm)",

@@

#png("../doc/display/census_precip.png", width=6, height=3, res=300, units="in") 
#png("../doc/display/census_precip.png", width=6, height=3, res=300, units="in") 
pre_yr_census
#dev.off()

@@

dendro.data<-dendrobyband %>%
  dplyr::rename(sp=SPCODE.UPDATE)%>%
  mutate(year=year(Date2)) %>%
  filter(is.na(DBH1_flag), DBH1_error=="none",
         win.decimal.error=="none", is.na(win.flag))%>%
  group_by(Cno, sp, Tag) %>%
  dplyr::summarise(calcDBH=calcDBH[which(n.dendro==min(n.dendro))],
                   Date2=Date2[which(n.dendro==min(n.dendro))],
                   Cii=max(Cii, na.rm=T))

@@

```
```{r long-term growth rates}

# remotes:::download("../data/HKK-dendro/stem1.RData",
#                    "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem1.rdata",
#                    auth_token = pat)
# remotes:::download("../data/HKK-dendro/stem2.RData",
#                    "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem2.rdata",
#                    auth_token = pat)
# remotes:::download("../data/HKK-dendro/stem3.RData",
#                    "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem3.rdata",
#                    auth_token = pat)
# remotes:::download("../data/HKK-dendro/stem4.RData",
#                    "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem4.rdata",
#                    auth_token = pat)

load("growth-precip-thailand/data/HKK-dendro/stem1.RData")
load("growth-precip-thailand/data/HKK-dendro/stem2.RData")
load("growth-precip-thailand/data/HKK-dendro/stem3.RData")
load("growth-precip-thailand/data/HKK-dendro/stem4.RData")

hkk_census <- hkk_census %>%
  filter(status=="A")%>%
  filter(tag %in% dendro.data$Tag)%>%
  group_by(tag, StemTag)%>%
  dplyr::mutate(dbhmin1 = dplyr::lag(dbh, n=1, default=NA, order_by=StemTag),
                inc_annual = (dbh-dbhmin1)/5)%>%
  filter(!is.na(dbhmin1))%>%
  #average increment in cm
  dplyr::summarise(avg_inc = mean(inc_annual/10, na.rm=T),
                   sd_inc = sd(inc_annual/10, na.rm=T))

#find the StemTags of trees that are in the dendro data
#match using the DBH04 column

#identify multistem trees from the census
multistem<-hkk_census %>%
  group_by(tag)%>%
  dplyr::summarise(n.stem=n())%>%
  rename(Tag=tag)

multistem<-merge(multistem, trees, by="Tag", all.x=T)

multistem<-multistem %>%
rename(tag=Tag)%>%
dplyr::mutate(dbh= (as.numeric(DBH.04)*10))

hkk.stem3.dendro<-hkk.stem3 %>%
  filter(tag %in% multistem$tag)

#add the StemTag values from hkk.stem3.dendro to multistem 
#match based on matching dbh and tag

multistem<- multistem %>%
dplyr::inner_join(select(hkk.stem3.dendro, c("tag", "StemTag", "dbh")), 
by=c("tag", "dbh"))

sum(multistem$tag != multistem$StemTag)
#there are no stemtags with values different from Tag
#all stems with dendrobands have them on the main stem?

#hkk_census<-hkk_census %>%
#  filter(StemTag %in% multistem$StemTag)

#since dendrobands are installed on trees with tag==StemTag
hkk_census<-hkk_census %>%
filter(tag==StemTag)

#there are 2271 trees in this dataset.

png("growth-precip-thailand/doc/display/census_increment.png", width=4, height=4, units="in", res=300)
inc_census
dev.off()

#plot average increment for the top 4 species

hkk_census2<-hkk_census %>%
dplyr::rename(Tag=tag)

hkk_census2<-merge(hkk_census2, trees, by="Tag")
hkk_census2<-hkk_census2 %>%
dplyr::rename(sp=SPCODE.UPDATE)

png("growth-precip-thailand/doc/display/census_increment_species.png", width=4, height=4, units="in", res=300)
ggplot(hkk_census2 %>% filter(sp %in% c("SACCLI", "HOPEOD", "TETRNU", "POLYVI")), 
aes(x= factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y = avg_inc, fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))) +
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
  scale_fill_viridis_d()+
  labs(title = 'tree growth in HKK from census', 
  y="Annualised diameter increment (cm)", x="species") +
  theme_bw()+
  #geom_vline(xintercept=0, lty=2)+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
dev.off()
```

```{r calculating increments for all years}

hkk_census<-hkk_census %>%
rename(Tag=tag)

#add long-term increment to dendro data
dendro.data.noouts<-merge(dendro.data.noouts, hkk_census, by="Tag", all.x=T)
dendro.data<-merge(dendro.data, hkk_census, by="Tag", all.x=T)

#making increments for each year

alt_census<-c(1, 3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 23, 25, 27, 29)

dendro_inc<- dendro.data %>%
  filter(Cno %in% alt_census)%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)
```
```{r species wise increments, echo=F}

#filter increments for two species

saccli<-dendro_inc %>%
  filter(sp == "SACCLI")

hopeod <- dendro_inc %>%
  filter(sp == "HOPEOD")

tetrnu<-dendro_inc %>%
  filter(sp == "TETRNU")

polyvi <- dendro_inc %>%
  filter(sp == "POLYVI")

#spaghetti plot saccli and hopeod increments over time

saccli_plot<-ggplot() +
  geom_line(data = saccli, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag)) +
  geom_line(data=saccli %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for SACCLI (deciduous)")+
  theme_bw()

hopeod_plot<-ggplot() +
  geom_line(data = hopeod, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag)) +
  geom_line(data=hopeod %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for HOPEOD (evergreen)")+
  theme_bw()

tetrnu_plot<-ggplot() +
  geom_line(data = tetrnu, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag)) +
  geom_line(data=tetrnu %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for TETRNU (deciduous)")+
  theme_bw()

polyvi_plot<-ggplot() +
  geom_line(data = polyvi, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag)) +
  geom_line(data=polyvi %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for POLYVI (evergreen)")+
  theme_bw()

png("growth-precip-thailand/doc/display/ind_increments.png", width=8, height=8, units="in", res=300)
grid.arrange(hopeod_plot, tetrnu_plot, polyvi_plot, saccli_plot, nrow=2)
dev.off()

top4<-dendro_inc %>%
filter(sp %in% c("SACCLI", "HOPEOD", "TETRNU", "POLYVI"))

#sample size per species per year
png("growth-precip-thailand/doc/display/top4_sample_size.png", width=8, height=4, units="in", res=300)
top4 %>% 
  group_by(sp, Cno) %>% 
  dplyr::summarise(n=n())%>%
  ggplot(aes(x=Cno, y=n, fill=factor(sp,levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_bar(stat="identity")+
  scale_fill_viridis_d()+
  facet_grid(.~factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))+
  xlab("census number")+
  ylab("sample size")+
  guides(fill=guide_legend(title="species"))+
  theme_bw()
dev.off()

png("growth-precip-thailand/doc/display/top4_increments.png", width=8, height=4, units="in", res=300)
ggplot(top4, aes(x = inc_annual, y = factor(((Cno-3)/2)+2009), 
                       fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "growth", option = "H", direction=-1) +
  labs(title = 'tree growth in HKK (cm)', y="year") +
  #theme_ipsum() +
  facet_grid(.~factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))+
  theme_bw()+
  xlim(-1.5, 1.5)+
  geom_vline(xintercept=0, lty=2)+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
dev.off()

growth_top4<-ggplot(top4, aes(x=factor(((Cno-3)/2)+2009), y=inc_annual,
                                              fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")),
                                              color=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))
                                              ))+ 
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
ylab("Annualised diameter increment (cm)")+theme_bw()+
facet_grid(.~factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))+
  geom_hline(yintercept = 0, lty=3)+
      theme(axis.text.x = element_text(angle=90))+
  ggtitle("HKK dendroband data from 2008 to 2023")+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  guides(fill=guide_legend(title="species"),
         color=guide_legend(title="species"))+
         xlab("year")

png("growth-precip-thailand/doc/display/top4_increments_violin.png", width=8, height=4, units="in", res=300)
growth_top4
dev.off()

growth_top4

growth_top4_sum<-ggplot(top4, aes(x=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y=inc_annual,
                                              fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")),
                                              color=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))
                                              ))+ 
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
ylab("Annualised diameter increment (cm)")+
xlab("species")+theme_bw()+
  geom_hline(yintercept = 0, lty=3)+
  ggtitle("tree growth in HKK from dendrobands")+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  guides(fill="none",
         color="none")

png("growth-precip-thailand/doc/display/top4_increments_sum.png", width=4, height=4, units="in", res=300)
growth_top4_sum
dev.off()

growth_top4
```

```{r plot increments}

dendro_2010<- dendro.data.noouts %>%
  filter(Cno %in% c(1, 3, 5))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

@@

library(gridExtra)
png("growth-precip-thailand/doc/display/increments_raw.png", width=8, height=4, units="in", res=300)
#grid.arrange(dendro_2010_plot, dendro_2015_plot, dendro_2019_plot, nrow=1)
grid.arrange(dendro_2010_plot, dendro_2015_plot, nrow=1)
dev.off()

@@

```{r top4 species drought resistance, echo=F}

resistance_2010_top4 <- top4 %>%
  filter(Cno %in% c(3, 5))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2015_top4 <- top4 %>%
  filter(Cno %in% c(11, 13, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

#calculate quantiles for resistance.div for both years

quantiles_2010_top4<-quantile(resistance_2010_top4$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)
quantiles_2015_top4<-quantile(resistance_2015_top4$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)

#split the data into categories based on quantiles

resistance_2010_top4$quantile<-ifelse(resistance_2010_top4$resistance.div<quantiles_2010_top4[1], 1, ifelse(resistance_2010_top4$resistance.div<quantiles_2010_top4[2], 2, ifelse(resistance_2010_top4$resistance.div<quantiles_2010_top4[3], 3, 4)))
resistance_2015_top4$quantile<-ifelse(resistance_2015_top4$resistance.div<quantiles_2015_top4[1], 1, ifelse(resistance_2015_top4$resistance.div<quantiles_2015_top4[2], 2, ifelse(resistance_2015_top4$resistance.div<quantiles_2015_top4[3], 3, 4)))

#map of resistance.div quantiles for top 4 species
resistance_all<-rbind(resistance_2010_top4, resistance_2015_top4)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010_top4)), rep(2015, nrow(resistance_2015_top4)))

resistance_map<-ggplot(resistance_all, aes(x=X, y=Y, size=as.numeric(DBH.04), col=quantile))+
  geom_point(alpha=0.5)+
   scale_size(range = c(.1, 10), name="DBH (cm)")+
  guides(color=guide_legend(title="resistance"))+
  scale_colour_gradient(low="red", high="blue") +
  facet_grid(sp~yr)+
  theme_minimal()+
  coord_fixed()

png("growth-precip-thailand/doc/display/resistance_map_top4.png", width=8, height=8, units="in", res=300)
resistance_map
dev.off()

@@

dec_williams<-read.csv("growth-precip-thailand/data/HKK-dendro/species.csv")

@@

resistance_2010 <- dendro_2010 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

@@

resistance_2010_div<-ggplot(data = resistance_2010, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  geom_vline(xintercept = 0, lty=2)+
  theme_bw()

@@

resistance_2015 <- dendro_2015 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

@@

resistance_2015_div<-ggplot(data = resistance_2015, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  geom_vline(xintercept = 0, lty=2)+
  theme_bw()

@@

library(gridExtra)

#png("growth-precip-thailand/doc/display/resistance_div.png", width=4, height=8, units="in", res=300)
#grid.arrange(resistance_2010_div, resistance_2015_div, resistance_2019_div, nrow=3)
#dev.off()

png("growth-precip-thailand/doc/display/resistance_div.png", width=8, height=4, units="in", res=300)
grid.arrange(resistance_2010_div, resistance_2015_div, nrow=1)
dev.off()

@@

quantiles_2010_div<-quantile(resistance_2010$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)

quantiles_2015_div<-quantile(resistance_2015$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)

quantiles_2019_div<-quantile(resistance_2019$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)

resistance_2010$quantile.div<-ifelse(resistance_2010$resistance.div<quantiles_2010_div[1], 1, ifelse(resistance_2010$resistance.div<quantiles_2010_div[2], 2,
            ifelse(resistance_2010$resistance.div<quantiles_2010_div[3], 3, 4))

resistance_2015$quantile.div<-ifelse(resistance_2015$resistance.div<quantiles_2015_div[1], 1, ifelse(resistance_2015$resistance.div<quantiles_2015_div[2], 2,
            ifelse(resistance_2015$resistance.div<quantiles_2015_div[3], 3, 4))

resistance_2019$quantile.div<-ifelse(resistance_2019$resistance.div<quantiles_2019_div[1], 1, ifelse(resistance_2019$resistance.div<quantiles_2019_div[2], 2,
            ifelse(resistance_2019$resistance.div<quantiles_2019_div[3], 3, 4))

resistance_all<-rbind(resistance_2010, resistance_2015)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010)), rep(2015, nrow(resistance_2015)))

@@

resistance_map<-ggplot(resistance_all, aes(x=X, y=Y, size=as.numeric(DBH.04), col=quantile.div))+
  geom_point(alpha=0.5)+
   scale_size(range = c(.1, 10), name="DBH (cm)")+
  guides(color=guide_legend(title="resistance"))+
  scale_colour_gradient(low="red", high="blue") +
  facet_grid(yr~.)+
  theme_minimal()+
  coord_fixed()

png("growth-precip-thailand/doc/display/resistance_map.png", width=8, height=6, units="in", res=300)
resistance_map
dev.off()
```
```{r basic model}

resistance_all$ciiF<-ordered(resistance_all$cii_min1)

model_2010<-glm(resistance.div ~ calcDBH_min1 + ciiF + williams_dec+
                  calcDBH_min1*ciiF + 
                  calcDBH_min1*williams_dec+
                  ciiF*williams_dec,
                data=resistance_all[resistance_all$yr==2010,])

@@

model_2015<-glm(resistance.div ~ calcDBH_min1 + ciiF + williams_dec+
                  calcDBH_min1*ciiF + 
                  calcDBH_min1*williams_dec+
                  ciiF*williams_dec,
                data=resistance_all[resistance_all$yr==2015,])

@@

@@

resistance_model<-resistance_all %>%
  select(resistance.div, calcDBH_min1, ciiF, williams_dec, yr, quantile)

@@

library(lme4)
library(nlme)
model_all<-lm(resistance.div ~ calcDBH_min1 + ciiF + williams_dec+
                  calcDBH_min1*ciiF + 
                  calcDBH_min1*williams_dec+
                  ciiF*williams_dec + yr,
                data=resistance_model,
                )

@@

plot_model(model_all, type="pred", terms=c("ciiF","williams_dec"), show.data = T)
plot_model(model_all, type="pred", terms=c("cii_min1"), show.data = T)
plot_model(model_all, type="pred", terms=c("williams_dec"), show.data = T)
plot_model(model_all, type="pred", terms=c("calcDBH_min1"), show.data = T)

@@

```{r data distribution, echo=F}

#growth data all years ridge plot
png("growth-precip-thailand/doc/display/growth_allyrs2.png", width=4, height=4, units="in", res=300)
ggplot(dendro_inc, aes(x = inc_annual, y = factor(((Cno-3)/2)+2009), 
                       fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "growth", option = "H", direction=-1) +
  labs(title = 'tree growth in HKK (cm)', y="year") +
  #theme_ipsum() +
  theme_bw()+
  xlim(-1.5, 1.5)+
  geom_vline(xintercept=0, lty=2)+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
dev.off()
